```{r}
library(ggplot2)
library(gridExtra)
library(grid)
library(randomForest)
library(party)
library(rpart)
```

```{r}
trainingData <- read.csv(file = 'train.csv',stringsAsFactors = FALSE,header = T)
dim(trainingData)
```

```{r}
names(trainingData)
```

```{r}
str(trainingData)
```

```{r}
head(trainingData,20)
```
```{r}
summary(trainingData)
```
The NA's : 177 shows that 177 values are missing in the Age column
```{r}
# replace the NA values with median age
trainingData[is.na(trainingData$Age),'Age'] <- median(trainingData$Age,na.rm = TRUE)
summary(trainingData)
```
```{r}
# visualize survival base on the gender
df <- aggregate(data.frame(trainingData$Sex,trainingData$Survived), by = list(trainingData$Sex, trainingData$Survived), FUN = length)
df <- data.frame(df[,1],df[,2],df[,3])
names(df) <- c("Sex","Survived","Count")
df[df[,"Survived"]==1,"Survived"] <- "Survived"
df[df[,"Survived"]==0,"Survived"] <- "Dead"
p <- ggplot(data=df,aes(x=df$Survived,y=df$Count,fill=df$Sex)) + geom_bar(stat = "identity")
p + labs(x = "Survived",y="Count", fill="Sex")
```



```{r}
survived <- trainingData$Survived
survived[survived==0] <- 'dead'
survived[survived==1] <- 'survived'
df <- data.frame(survived,trainingData$Age)
names(df) <- c("Survived","Age")
ggplot(df, aes(x=Age,fill=Survived)) + geom_histogram(binwidth = 2) 
```

```{r}
# Let's now focus on the Fare ticket of each passenger and correlate it with the survival
survived <- trainingData$Survived
survived[survived==0] <- 'dead'
survived[survived==1] <- 'survived'
df <- data.frame(survived,trainingData$Fare)
names(df) <- c("Survived","Fare")
ggplot(df, aes(x=Fare,fill=Survived)) + geom_histogram(binwidth = 10) 
```
Passengers with cheaper ticket fares are more likely to die. Put differently, passengers with more expensive tickets, and therefore a more important social status, seem to be rescued first.
```{r}
# Let's now combine the age, the fare and the survival on a single chart.
p <- ggplot(trainingData,aes(trainingData$Age,trainingData$Fare)) + geom_point(aes(colour = survived))
survived <- factor(trainingData$Survived)
p + labs(x = "Age",y="Fare")
```
A distinct cluster of dead passengers (the red one) appears on the chart. Those people are adults (age between 15 and 50) of lower class (lowest ticket fares).

```{r}
agg <- aggregate(trainingData$Fare~trainingData$Pclass, trainingData,mean)
names(agg) <- c("Pclass","Fare")
ggplot(agg, aes(agg$Pclass, agg$Fare)) + geom_col(colour="tomato1",fill="#FF9999") + labs(x = "Pclass",y="Fare")


```

```{r}
# Let's now see how the embarkation site affects the survival.
trainingData[trainingData$Embarked=='',"Embarked"] <- 'S'
df <- aggregate(data.frame(trainingData$Embarked,trainingData$Survived), by = list(trainingData$Embarked, trainingData$Survived), FUN = length)
df <- data.frame(df[,1],df[,2],df[,3])
names(df) <- c("Embarked","Survived","Count")
df[df[,"Survived"]==1,"Survived"] <- "Survived"
df[df[,"Survived"]==0,"Survived"] <- "Dead"
ggplot(data=df,aes(x=Survived,y=Count,fill=Embarked)) + geom_bar(stat = "identity")
```
There seems to be no distinct correlation here
```{r}
# combine training data and testing data together
getCombinedData <- function(){
  trainingData <- read.csv(file = 'train.csv',stringsAsFactors = FALSE,header = T)
  testingData <- read.csv(file = 'test.csv',stringsAsFactors = FALSE,header = T)
  testingData$Survived <- NA
  combined <- rbind(trainingData,testingData)
  return(combined)
}
combined <- getCombinedData()
head(combined)
```

```{r}
# extracting the passenger titles
# If you look closely at these first examples:
# Braund, Mr. Owen Harris
# Heikkinen, Miss. Laina
# Oliva y Ocana, Dona. Fermina
# Peter, Master. Michael J
# You will notice that each name has a title in it ! This can be a simple Miss. or Mrs. but it can be sometimes something more sophisticated like Master, Sir or Dona. In that case, we might introduce an additional information about the social status by simply parsing the name and extracting the title.

getTitles <- function(){
  # ",Capt\\.", "\\." means escape sequence. It can be converted to ",Capt.",detect which name contains this string and convert it to "Officer"
  titleDictionary = c("Officer","Officer","Officer","Royalty","Royalty","Royalty","Officer","Officer","Royalty",
                      "Royalty","Mrs","Miss","Mrs","Mr","Mrs","Miss","Master","Royalty")
  names(titleDictionary) = c(", Capt\\.",", Col\\.",", Major\\.",", Jonkheer\\.",", Don\\.",", Sir\\.",", Dr\\.",", Rev\\.",
                             ", the Countess\\.",", Dona\\.",", Mme\\.",", Mlle\\.",", Ms\\.",", Mr\\.",", Mrs\\.",", Miss\\.",
                             ", Master\\.",", Lady\\.")
  
  return(titleDictionary)
}
titles <- getTitles()
# convert Name column to Title
convertNameToTitle <- function(titles){
  allNames = combined$Name
  for (i in 1:length(names(titles))){
    index = grep(names(titles)[i],allNames)
    allNames[index] = titles[i]
  }
  
  return(allNames)
}
titleColumn <- convertNameToTitle(titles)
# combined$Title <- titleColumn
combined <- cbind(combined[,1:4],Title = titleColumn,combined[,5:ncol(combined)])
# use the function above, all female title will be converted to Mr too.
table(titleColumn)

```

```{r}
# create family size feature

table(combined$Parch)
getFamilySize <- function(row){
  c(row[1] + row[2])
}
familySize <- apply(combined[,c("Parch","SibSp")],1,getFamilySize)
combined$FamilySize <- familySize
```

```{r}
# analize the age column
df <- data.frame(combined$Sex,combined$Pclass,combined$Title,combined$PassengerId,combined$Age,combined$SibSp,combined$Parch,combined$Fare)
names(df) <- c("Sex","Pclass","Title","PassengerId","Age","SibSp","Parch","Fare")
df$Sex <- as.character(df$Sex)
df$Title <- as.character(df$Title)
df <- aggregate(data.frame(df$PassengerId,df$Age,df$SibSp,df$Parch,df$Fare),by = list(df$Title,df$Pclass, df$Sex), FUN = median,na.rm=TRUE)
names(df) <- c("Title","Pclass","Sex","PassengerId","Age","SibSp","Parch","Fare")
df <- df[,c("Sex","Pclass","Title","PassengerId","Age","SibSp","Parch","Fare")]
df
```

```{r}
# use decision tree to fill missing ages
Agefit <- rpart(Age ~ Pclass + Sex + SibSp + Parch + Fare + Embarked + Title,
                  data=combined[!is.na(combined$Age),], 
                  method="anova")
combined$Age[is.na(combined$Age)] <- predict(Agefit, combined[is.na(combined$Age),])
head(combined,20)
```

```{r}
# create a function and get missing age base on the features
# getAge <- function(row){
#   if(row["Sex"]=="female"){
#     if(row["Pclass"]==1){
#       if(row["Title"]=="Miss"){
#         return (df[df[,"Sex"]=="female"&df[,"Pclass"]==1&df[,"Title"]=="Miss","Age"])
#       }else if(row["Title"]=="Mrs"){
#         return (df[df[,"Sex"]=="female"&df[,"Pclass"]==1&df[,"Title"]=="Mrs","Age"])
#       }else if(row["Title"]=="Officer"){
#         return (df[df[,"Sex"]=="female"&df[,"Pclass"]==1&df[,"Title"]=="Officer","Age"])
#       }else if(row["Title"]=="Royalty"){
#         return (df[df[,"Sex"]=="female"&df[,"Pclass"]==1&df[,"Title"]=="Royalty","Age"])
#       }else{
#         print("something wrong")
#       }
#     }else if(row["Pclass"]==2){
#       if(row["Title"]=="Miss"){
#         return (df[df[,"Sex"]=="female"&df[,"Pclass"]==2&df[,"Title"]=="Miss","Age"])
#       }else if(row["Title"]=="Mrs"){
#         return (df[df[,"Sex"]=="female"&df[,"Pclass"]==2&df[,"Title"]=="Mrs","Age"])
#       }else{
#         print("something wrong")
#       }
#     }else if(row["Pclass"]==3){
#       if(row["Title"]=="Miss"){
#         return (df[df[,"Sex"]=="female"&df[,"Pclass"]==3&df[,"Title"]=="Miss","Age"])
#       }else if(row["Title"]=="Mrs"){
#         return (df[df[,"Sex"]=="female"&df[,"Pclass"]==3&df[,"Title"]=="Mrs","Age"])
#       }else{
#         print("something wrong")
#       }
#     }else{
#       print("something wrong")
#     }
#   }else if(row["Sex"]=="male"){
#     if(row["Pclass"]==1){
#       if(row["Title"]=="Master"){
#         return (df[df[,"Sex"]=="male"&df[,"Pclass"]==1&df[,"Title"]=="Master","Age"])
#       }else if(row["Title"]=="Mr"){
#         return (df[df[,"Sex"]=="male"&df[,"Pclass"]==1&df[,"Title"]=="Mr","Age"])
#       }else if(row["Title"]=="Officer"){
#         return (df[df[,"Sex"]=="male"&df[,"Pclass"]==1&df[,"Title"]=="Officer","Age"])
#       }else if(row["Title"]=="Royalty"){
#         return (df[df[,"Sex"]=="male"&df[,"Pclass"]==1&df[,"Title"]=="Royalty","Age"])
#       }else{
#         print("something wrong")
#       }
#     }else if(row["Pclass"]==2){
#       if(row["Title"]=="Master"){
#         return (df[df[,"Sex"]=="male"&df[,"Pclass"]==2&df[,"Title"]=="Master","Age"])
#       }else if(row["Title"]=="Mr"){
#         return (df[df[,"Sex"]=="male"&df[,"Pclass"]==2&df[,"Title"]=="Mr","Age"])
#       }else if(row["Title"]=="Officer"){
#         return (df[df[,"Sex"]=="male"&df[,"Pclass"]==2&df[,"Title"]=="Officer","Age"])
#       }else{
#         print("something wrong")
#       }
#     }else if(row["Pclass"]==3){
#       if(row["Title"]=="Master"){
#         return (df[df[,"Sex"]=="male"&df[,"Pclass"]==3&df[,"Title"]=="Master","Age"])
#       }else if(row["Title"]=="Mr"){
#         return (df[df[,"Sex"]=="male"&df[,"Pclass"]==3&df[,"Title"]=="Mr","Age"])
#       }else{
#         print("something wrong")
#       }
#     }else{
#       print("something wrong")
#     }
#   }else{
#     print("something wrong")
#   }
# }
# 
# missingAges <- apply(combined[is.na(combined[,"Age"]),],1,getAge)
# combined[is.na(combined[,"Age"]),"Age"] <- missingAges
# head(combined)
```

```{r}
# fill missing fare
boxplot(combined$Fare,ylab = 'Fares',main='Distribution of Fares',col='red')
upper.whisker <- boxplot.stats(combined$Fare)$stats[5]
outlier.filter <- combined$Fare < upper.whisker
fare.equation = "Fare ~ Pclass + Sex + Age + SibSp + Parch + Embarked"
fare.model <- lm(
  formula = fare.equation,
  data = combined[outlier.filter,]
)

missingFare.row <- combined[
  is.na(combined$Fare),
  c("Pclass","Sex","Age","SibSp","Parch","Embarked")
  ]
fare.predictions <- predict(fare.model,missingFare.row)
fare.predictions
combined[is.na(combined$Fare),"Fare"] <- fare.predictions

# combined$Fare[1044] <- median(combined$Fare, na.rm=TRUE)
combined$Fare[1044]
```

```{r}
# fill missing embarked with most frequent embarked value
table(combined$Embarked)
combined[combined[,"Embarked"]=="","Embarked"] <- 'S'
```

```{r}
####################### use random forest to predict the survived column of testing data  ####################
# categorical casting
combined$Sex <- as.factor(combined$Sex)
combined$Cabin <- as.factor(combined$Cabin)
combined$Embarked <- as.factor(combined$Embarked)

# read back training data and testing data from full data
titanic.train <- combined[!is.na(combined$Survived),]
titanic.test <- combined[is.na(combined$Survived),]

titanic.train$Survived <- as.factor(titanic.train$Survived)

survived.equation <- "Survived ~ Age + Fare + Sex + Title + Pclass + SibSp + Parch + Embarked + FamilySize"
survived.formula <- as.formula(survived.equation)


# construct prediction model
titanic.model <- randomForest(formula = survived.formula,data=titanic.train, ntree = 2500, mtry = 3, importance = TRUE)

# check which feature is important
varImpPlot(titanic.model)

Survived <- predict(titanic.model,newdata = titanic.test)

PassengerId <- titanic.test$PassengerId
output.df <- data.frame(PassengerId,Survived)
# write.csv(output.df, file = "kaggle_submission8.csv", row.names = FALSE)
```

```{r}
###################### use conditional inference trees
# set.seed(312)
# # categorical casting
# combined$Sex <- as.factor(combined$Sex)
# combined$Cabin <- as.factor(combined$Cabin)
# combined$Embarked <- as.factor(combined$Embarked)
# combined$Survived <- as.factor(combined$Survived)
# # read back training data and testing data from full data
# titanic.train <- combined[!is.na(combined$Survived),]
# titanic.test <- combined[is.na(combined$Survived),]
# 
# fit <- cforest(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare +
#                                        Embarked + Title + FamilySize,
#                  data = titanic.train, 
#                  controls=cforest_unbiased(ntree=2500, mtry=3))
# 
# Prediction <- predict(fit, titanic.test, OOB=TRUE, type = "response")
# output.df <- data.frame(PassengerId = titanic.test$PassengerId,Survived = Prediction)
# write.csv(output.df, file = "kaggle_submission7.csv", row.names = FALSE)
```

